Co-citations in context

### 2019-06-04 cocit_r.tex
Start of material rearrangement

\section{Materials and Methods}
\emph{Bibliographic data} We have previously developed ERNIE, an open source knowledge platform into which we parse the Web of Science (WoS) Core Collection~\citep{Keserci371955}. WoS data stored in ERNIE spans the period 1900-2019 and consists of over 72 million publications. For this study, we generated an analytical dataset from years 1985 to 2005 from ERNIE. The total number of publications in this dataset was just over 25 million publications (25,134,073).  For each of these years, we further restricted analysis to publications of type Article. Since WoS data also contains incomplete references or references that point at other indexes, we also considered only those references for which there were complete records~(\ref{tab:summary_data}). For example, WoS data for year 2005 contained 1,753,174 publications, which after restricting to type Article and considering only those references described above resulted in 916,573 publications, 6,095,594 unique references (set of references), and 17,167,347 total references (multiset of references). Given consistent trends in the data, we analyzed the two boundary years (1985 and 2005) and the mid-point (1995) performing 1,000 simulations for each dataset.

\emph{Disciplinary datasets} We constructed three disciplinary datasets based on keyword searches. (i) immunology (ii) metabolism (iii) applied physics. For the first two, rooted in biomedical research, we searched Pubmed for the term `immunology' or `metabolism' in the years 1985, 1995, and 2005~(\ref{tab:disc-datasets}). Pubmed IDs (pmids) returned were matched to WoS IDs (wos\_ids) and used to retrieve relevant articles. For the applied physics dataset, we directly searched subject labels in WoS for `applied physics'. We also examined publications in the five major research areas in the Web of Science; life sciences \& biomedicine, physical sciences, technology, social sciences, arts \& humanities using the extended subcategory classification of 153 sub-groups to categorize disciplinary composition of cited references in the datasets we studied.

@@

\subsection{Chacko} In our study, we also use a Monte Carlo approach to simulate under a random graph model. A  principal consideration, however, was to restrict model misspecification arising from disciplinarily irrelevant references. We addressed this consideration by analyzing disciplinary subsets of the scientific literature, thereby restricting random selection of references to only those references in the disciplinary network being studied. We predicted that model misspecification as measured by the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and simulated frequencies in a disciplinary network would be less than the divergence for observed and simulated frequencies in the WoS superset. The results indicate that for the set of journals common to both a disciplinary network and the WoS superset~\ref{tab:KLD}, simulations under our model consistently have a lower K-L divergence compared to simulations that draw from the WoS superset (and its attendant substitutions that are ectopic with respect to field and discipline).

\begin{figure}
\centering
\includegraphics[width=0.5\hsize]{fold_effect.pdf}
\caption{Intra-network citation shuffling preserves disciplinary composition. Publications of type Article belonging to the WoS research areas (i) Arts and Humanities (ii) Life Sciences and Biomedicine were subject to a single shuffle of all their cited references using either the cited references in these networks as a source of random substitutions (bg\_local) or references from all articles in WoS (bg\_allWoS). The disciplinary composition of cited reference before and after shuffling was measured for each of 151 sub-groups (from the extended subject classification in WoS) in these two research areas, expressed as fold difference between citation counts grouped by subject for original (o) and shuffled (s) references using the formula (fold\_difference = $ifelse(o > s, o/s, s/o)$) and rounded to the nearest integer. A fold difference of 1 indicates that citation shuffling did not alter disciplinary composition. Data are shown for articles published in 1985. All four boxplots are generated from 151 observations. Min \& Max values for fold differences were as follows. (i) Arts \& Humanities. bg\_local (1,2), bg\_allWoS (1, 488) (ii) Life Sciences \& Biomedicine. bg\_local(1,1), bg\_allWos (1, 352) Note y-axis: $log_2$ scale.} 
\end{figure}

\begin{figure}
\includegraphics[width=\hsize]{Fig1}
\caption{Fraction of publications with high conventionality (HC) and high novelty (HN) signatures relative to citation count.}
\end{figure}

\subsection{Bradley}
Figure \ref{fig:scatter} shows that the z-scores for the same journal pair can be positive (negative) when computed with respect to one data set but be negative (positive) for another data set. The journal-pair z-scores in Figure \ref{fig:scatter} have consistent signs for both Immunology and WoS data sets in 71.4\% of the instances and different signs for 28.6\% of the journal pairs. A negative journal pair z-score in one context increases the likelihood that articles citing it will be deemed to be novel while articles citing it in another context, where the journal pair z-score is positive, are less likely to be classified as novel and possibly more likely to be classified as conventional. Figure \ref{fig:scatter} reflects that the WoS data set has approximately 44,000 fewer negative z-scores than does the immunology data set, which contributes to its significantly lower percentage of high-novelty articles as shown in Figure \ref{fig:scatter}.

@@

>>>>

\section{Materials and Methods}
\emph{Bibliographic data} We have previously developed ERNIE, an open source knowledge platform into which we parse the Web of Science (WoS) Core Collection~\citep{Keserci371955}. WoS data stored in ERNIE spans the period 1900-2019 and consists of over 72 million publications. For this study, we generated an analytical dataset from years 1985 to 2005 from ERNIE. The total number of publications in this dataset was just over 25 million publications (25,134,073).  For each of these years, we further restricted analysis to publications of type Article. Since WoS data also contains incomplete references or references that point at other indexes, we also considered only those references for which there were complete records~(Table \ref{tab:summary_data}). For example, WoS data for year 2005 contained 1,753,174 publications, which after restricting to type Article and considering only those references described above resulted in 916,573 publications, 6,095,594 unique references (set of references), and 17,167,347 total references (multiset of references). Given consistent trends in the data, we analyzed the two boundary years (1985 and 2005) and the mid-point (1995) performing 1,000 simulations for each dataset.

\emph{Disciplinary datasets} We constructed three disciplinary datasets based on keyword searches. (i) immunology (ii) metabolism (iii) applied physics. For the first two, rooted in biomedical research, we searched Pubmed for the term `immunology' or `metabolism' in the years 1985, 1995, and 2005~(Table \ref{tab:disc-datasets}). Pubmed IDs (pmids) returned were matched to WoS IDs (wos\_ids) and used to retrieve relevant articles. For the applied physics dataset, we directly searched subject labels in WoS for `applied physics'. We also examined publications in the five major research areas in the Web of Science; life sciences \& biomedicine, physical sciences, technology, social sciences, arts \& humanities using the extended subcategory classification of 153 sub-groups to categorize disciplinary composition of cited references in the datasets we studied.

@@

\emph{This paragraph needs work, but I wanted to create a placeholder for the introduction to this subsection} A central theme of this article is the misspecification of the model in Uzzi et al. in terms of its sampling from a broad dataset without consideration of observed citation practices where papers cite a high percentage of papers from within the discipline related to the paper, frequency of citation, and xxx. For example we found (\emph{cite the statistics from Sitaram's study and other references about citation practices.}) We demonstrate the effects of model misspecification and effectiveness of using disciplinary datasets to resolve those issues.

Figure \ref{fig:scatter} shows that the z-scores for the same journal pair can be positive (negative) when computed with respect to one data set but be negative (positive) for another data set. The journal-pair z-scores in Figure \ref{fig:scatter} have consistent signs for both Immunology and WoS data sets in 71.4\% of the instances and different signs for 28.6\% of the journal pairs. When a journal-pair z-score is negative with respect to one dataset and positive with respect to another dataset, articles citing that pair are more likely to be deemed novel in the first instance and less likely to be deemed novel in the second instance.  It seems that any journal pair should either be indicative of novelty or not, but it should never have contradictory implications with a changing reference dataset. We view, therefore, these contradictions as symptomatic inappropriately sampling from a broad dataset and ignoring observed citation patterns.  Figure \ref{fig:scatter} reflects that the WoS data set has approximately 44,000 fewer negative z-scores than does the immunology data set, which contributes to its significantly lower percentage of high-novelty articles.  \emph{The plots I made of the cumulative z-score distributions for Immunology and WoS also could be used to demonstrate this phenomenon, although the scatter plot does this more clearly, and it is more striking.  The scatter plot also encodes more data, that is, the matched z-scores for each journal pair.}

Note also in Figure \ref{fig:scatter} that many z-scores change values significantly between the WoS and Immunology datasets, although the density of points near the origin in Figure \ref{fig:scatter} is significant where the z-scores may have different signs with respect to the two datasets, although their magnitudes are not significantly different. While no metric is without downside, this observation points to a weakness of measuring novelty based on the sign of the z-score where the implications of a small negative value and a large negative value are the same.

@@

The variation in z-scores with respect to the reference datasets causes the percentages of articles denoted as highly conventional and highly novel to be significantly different, as demonstrated in Figure \ref{fig:Fig1}. \emph{Insert text about Figure \ref{fig:Fig1}.}

\begin{figure}
\includegraphics[width=\hsize]{Fig1}
\caption{Fraction of publications with high conventionality (HC) and high novelty (HN) signatures relative to citation count.      \emph{What follows is the caption from PNAS submission: is any of this useful?}  Effect of Research Discipline,  Background Network,  and Citation Count on Conventionality and Novelty.
Data are shown for the applied physics (18,305), immunology (21,917), metabolism (97,405) and WoS (476,288) networks for 1995. The number of publications in each network is shown in parentheses. Citation counts shown are cumulative over the first 8 years since publication. 
X-axis: publications were classified into percentile groups based on citation counts (e.g., Top 1 indicates those publications in the top 1\%). Y axis: The percent of applications in each group that are high conventionality (HC) and high novelty (HN). The z-scores are computed for each disciplinary network based on the selected background network; thus, \emph{imm} denotes the immunology network with immunology z-scores and \emph{imm\_wos} denotes the immunology network with z-scores from WoS z-scores. The figure shows striking differences between the WOS network compared to the metabolism and immunology networks: across all networks, the percentage of high conventionality (HC) publications increases with citation counts, while the percentage of high novelty (HN) publications decreases with citation counts for the biological networks but not for WOS. \label{fig:Fig1}}
\end{figure}

Another possible symptom of model misspecification is to find that the citations associated with an article under random sampling do not match with the observed disciplinary makeup of citation in practice. \emph{More needed here}  Here is Sitaram's study:

Adding to the point of papers citing papers belonging to the same discipline, I collected the subject areas for all the cited references from the 1985 Applied Physics dataset and searched for the keyword `Physics`. Then calculated the percentage of references related to Applied Physics for each paper References-related-to-physics/Total-references. Only papers with at-least 5 references were considered and the total publications were 6991.
Results show a mean of 0.77 i.e. on an average papers cite 77\% of their references belonging to same discipline.

\noindent count    6991.000000\\
mean        0.774002

We investigated the effect of sampling from disciplinary data sets versus more broadly from the WoS on the disciplinary make-up of the citations in articles under random sampling (\emph{sentence needs work but is an example of a transition.}) \emph{Insert text about Figure \ref{fold_effect}, specifically the right-hand side of Figure  \ref{fold_effect} where the fold statistics demonstrate the effect with Uzzi's approach.}

Like Uzzi et al., we also used a Monte Carlo approach to simulate under a random graph model. A  principal consideration, however, was to restrict model misspecification arising from disciplinarily irrelevant references. We addressed this consideration by analyzing disciplinary subsets of the scientific literature, thereby restricting random selection of references to only those references in the disciplinary network being studied.  \emph{Insert comment about the left-hand side of Figure \ref{fold_effect} where the disciplinary dataset has resolved the fold issue.}  We also predicted that model misspecification as measured by the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and simulated frequencies in a disciplinary network would be less than the divergence for observed and simulated frequencies in the WoS superset. The results in Table \ref{tab:KLD} indicate that for the set of journals common to both a disciplinary network and the WoS superset, simulations under our model consistently have a lower K-L divergence compared to simulations that draw from the WoS superset (and its attendant substitutions that are ectopic with respect to field and discipline).

\begin{figure}
\label{fold_effect}
\centering
\includegraphics[width=0.5\hsize]{fold_effect.pdf}
\caption{Intra-network citation shuffling preserves disciplinary composition. Publications of type Article belonging to the WoS research areas (i) Arts and Humanities (ii) Life Sciences and Biomedicine were subject to a single shuffle of all their cited references using either the cited references in these networks as a source of random substitutions (bg\_local) or references from all articles in WoS (bg\_allWoS). The disciplinary composition of cited reference before and after shuffling was measured for each of 151 sub-groups (from the extended subject classification in WoS) in these two research areas, expressed as fold difference between citation counts grouped by subject for original (o) and shuffled (s) references using the formula (fold\_difference = $ifelse(o > s, o/s, s/o)$) and rounded to the nearest integer. A fold difference of 1 indicates that citation shuffling did not alter disciplinary composition. Data are shown for articles published in 1985. All four boxplots are generated from 151 observations. Min \& Max values for fold differences were as follows. (i) Arts \& Humanities. bg\_local (1,2), bg\_allWoS (1, 488) (ii) Life Sciences \& Biomedicine. bg\_local(1,1), bg\_allWos (1, 352) Note y-axis: $log_2$ scale.} 
\end{figure}
